Model Comparisons

### 2015-11-29 models/Model_Comparison.Rmd
plots with both e6 and e11 data for comparison - plot with combined prag listener judgments?

# 3.c)
# Emp alts
# Normalized = T
# alpha tune = T
# compare.data = Toggle this with data.partial for e6 or NULL for e11
alphas[11] = tune.alpha(data.full, type = "full", compare.data=data.partial, addMid=F, normalize=T)
m3.c = data.full %>%
  do(run.full(., alpha = alphas[11], addMid=F, normalize=T))
# Match with e6 w/ checks
m3.c.matched = m3.c[matched.items, ]
m3.c.matched$degree = ifelse(m3.c.matched$degree == "hi1", "hi", "low")
m3.c.matched = cbind(m3.c.matched, data.partial$listener.p)
all(m3.c.matched$scale == data.partial$scale & m3.c.matched$degree == data.partial$degree)
colnames(m3.c.matched)[length(colnames(m3.c.matched))] = "e6.listener.p"
# Store output
performance.output[11, ] = c("m3.c", round(cor(m3.c.matched$e6.listener.p, m3.c.matched$pred), 5),
                             round(cor(m3.c.matched$listener.p, m3.c.matched$pred), 5), T)
# Store plot
m3.c.plot = qplot(stars, e6.listener.p, col=degree, 
      data=m3.c.matched,
      main=paste("m3.c\nalpha: ", alphas[11],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[11, 2]),
      ylab="Posterior p(rating | word)") +
      ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

@@

@@

# Store plot
m4.c.plot = qplot(stars, listener.p, col=degree, 
      data=m4.c.matched,
      main=paste("m4.c\nalpha: ", alphas[15],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[15, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 4.d)
# Emp alts + extras
# Normalized = F
# alpha tune = T
# compare.data = Toggle this with data.partial or NULL
alphas[16] = tune.alpha(data.full.extras, type = "full", compare.data=data.partial, addMid=T, normalize=F)
m4.d = data.full.extras %>%
  do(run.full(., alpha = alphas[16], addMid=T, normalize=F))
# Match with e6 w/ checks
m4.d.matched = m4.d[matched.items.extras, ]
m4.d.matched$degree = ifelse(m4.d.matched$degree == "hi1", "hi", "low")
m4.d.matched = cbind(m4.d.matched, data.partial$listener.p)
all(m4.d.matched$scale == data.partial$scale & m4.d.matched$degree == data.partial$degree)
colnames(m4.d.matched)[length(colnames(m4.d.matched))] = "e6.listener.p"
# Store output
performance.output[16, ] = c("m4.d", round(cor(m4.d.matched$e6.listener.p, m4.d.matched$pred), 5),
                             round(cor(m4.d.matched$listener.p, m4.d.matched$pred), 5), F)
# Store plot
m4.d.plot = qplot(stars, e6.listener.p, col=degree, 
      data=m4.d.matched,
      main=paste("m4.d\nalpha: ", alphas[16],
                 "\nNormalized = ", F,
                 "\nCorr = ", performance.output[16, 2]),
      ylab="Posterior p(rating | word)") +
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)
```

# Plots

@@

`m2:` Entailment + generic None models (i.e. "some_all" scale includes "some", "all", "none")

`m3:` Full model with empirical alternatives (i.e. "some_all" scale includes "some", "all", "most", "none")

`m4:` Entailment + empirical alternatives + additional fitted scalar (i.e. "some_all" scale includes "some", "all", "most", "none", "mid")

@@

Initial thoughts:

We see fairly steady model improvement with the addition of salient alternatives. Let's focus on the normalized family of models (1a, 1c, ... , 4a, 4c) - I explain why this might be a good idea below, but I'm not totally convinced. Probably the most important comparisons to pay attention to is the jump in correlation values when we first introduce empirically derived alternatives (models m3). Comparing m2 models (entailment + generic None) and m3 models (empirically derived alternatives) we see jumps of +0.23 for r values with alpha tuning. See the plots below:
```{r}
m2.c.plot
m3.c.plot
```
Importantly, we also see improvements with the addition of 'fitted' alternatives in m4 see the plots below:
```{r}
m3.c.plot
m4.c.plot
```

Additional notes:

1) In general model performance is greater (correlation) for e11 data. This may be due in part to less extreme literal listener evaluations in e11 (see RSA+AmbiguousSemantics for a more detailed analysis for literal listener experiment differences). Our best model performance occurs with e11 and the model is m4.c, correlation = `r round(performance.outpu[15, 3], 5)`.

@@

3) We may want to explore the the idea of unique alphas for a given scale (would need to change how we've written RSA), for example using scale entropy measures.
```{r overall_perf}
performance.output = cbind(performance.output, alphas)
grid.table(performance.output)

m1.a.plot
m1.b.plot
m1.c.plot
m1.d.plot

>>>>

# 3.c)
# Emp alts
# Normalized = T
# alpha tune = T
# compare.data = Toggle this with data.partial for e6 or NULL for e11
alphas[11] = tune.alpha(data.full, type = "full", compare.data=data.partial, addMid=F, normalize=T)
m3.c = data.full %>%
  do(run.full(., alpha = alphas[11], addMid=F, normalize=T))
# Match with e6 w/ checks
m3.c.matched = m3.c[matched.items, ]
m3.c.matched$degree = ifelse(m3.c.matched$degree == "hi1", "hi", "low")
m3.c.matched = cbind(m3.c.matched, data.partial$listener.p)
all(m3.c.matched$scale == data.partial$scale & m3.c.matched$degree == data.partial$degree)
colnames(m3.c.matched)[length(colnames(m3.c.matched))] = "e6.listener.p"
# Store output
performance.output[11, ] = c("m3.c", round(cor(m3.c.matched$e6.listener.p, m3.c.matched$pred), 5),
                             round(cor(m3.c.matched$listener.p, m3.c.matched$pred), 5), T)
# Store plot 1
m3.c.plot = qplot(stars, e6.listener.p, col=degree, 
      data=m3.c.matched,
      main=paste("m3.c\nalpha: ", alphas[11],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[11, 2]),
      ylab="Posterior p(rating | word)") +
      ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# Store plot 2 - both e6 and e11 data
m3.c.jointPlot = m3.c.matched %>%
  gather(study, listener.pred, listener.p, e6.listener.p)
m3.c.jointPlot$study = ifelse(m3.c.jointPlot$study == "e6.listener.p", "e6", "e11")
qplot(stars, listener.pred, col=degree, shape=study,
      data=m3.c.jointPlot,
      main=paste("m3.c\nalpha: ", alphas[11],
                 "\nNormalized = ", T,
                 ylab="Posterior p(rating | word)")) +
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

@@

# Store plot 2 - both e6 and e11 data
m3.d.jointPlot = m3.d.matched %>%
  gather(study, listener.pred, listener.p, e6.listener.p)
m3.d.jointPlot$study = ifelse(m3.d.jointPlot$study == "e6.listener.p", "e6", "e11")
qplot(stars, listener.pred, col=degree, shape=study,
      data=m3.d.jointPlot,
      main=paste("m3.d\nalpha: ", alphas[12],
                 "\nNormalized = ", T,
                 ylab="Posterior p(rating | word)")) +
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

@@

# Store plot
m4.c.plot = qplot(stars, e6.listener.p, col=degree, 
      data=m4.c.matched,
      main=paste("m4.c\nalpha: ", alphas[15],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[15, 2]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# Store plot 2 - both e6 and e11 data
m4.c.jointPlot = m4.c.matched %>%
  gather(study, listener.pred, listener.p, e6.listener.p)
m4.c.jointPlot$study = ifelse(m4.c.jointPlot$study == "e6.listener.p", "e6", "e11")
qplot(stars, listener.pred, col=degree, shape=study,
      data=m4.c.jointPlot,
      main=paste("m4.c\nalpha: ", alphas[15],
                 "\nNormalized = ", T,
                 ylab="Posterior p(rating | word)")) +
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# 4.d)
# Emp alts + extras
# Normalized = F
# alpha tune = T
# compare.data = Toggle this with data.partial or NULL
alphas[16] = tune.alpha(data.full.extras, type = "full", compare.data=data.partial, addMid=T, normalize=F)
m4.d = data.full.extras %>%
  do(run.full(., alpha = alphas[16], addMid=T, normalize=F))
# Match with e6 w/ checks
m4.d.matched = m4.d[matched.items.extras, ]
m4.d.matched$degree = ifelse(m4.d.matched$degree == "hi1", "hi", "low")
m4.d.matched = cbind(m4.d.matched, data.partial$listener.p)
all(m4.d.matched$scale == data.partial$scale & m4.d.matched$degree == data.partial$degree)
colnames(m4.d.matched)[length(colnames(m4.d.matched))] = "e6.listener.p"
# Store output
performance.output[16, ] = c("m4.d", round(cor(m4.d.matched$e6.listener.p, m4.d.matched$pred), 5),
                             round(cor(m4.d.matched$listener.p, m4.d.matched$pred), 5), F)
# Store plot
m4.d.plot = qplot(stars, e6.listener.p, col=degree, 
      data=m4.d.matched,
      main=paste("m4.d\nalpha: ", alphas[16],
                 "\nNormalized = ", F,
                 "\nCorr = ", performance.output[16, 2]),
      ylab="Posterior p(rating | word)") +
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

# Store plot 2 - both e6 and e11 data
m4.d.jointPlot = m4.d.matched %>%
  gather(study, listener.pred, listener.p, e6.listener.p)
m4.d.jointPlot$study = ifelse(m4.d.jointPlot$study == "e6.listener.p", "e6", "e11")
qplot(stars, listener.pred, col=degree, shape=study,
      data=m4.d.jointPlot,
      main=paste("m4.d\nalpha: ", alphas[16],
                 "\nNormalized = ", T,
                 ylab="Posterior p(rating | word)")) +
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)
```

# Discussion

@@

`m2:` Entailment + generic None models (i.e. "some_all" scale includes "none", "some", "all")

`m3:` Full model with empirical alternatives (i.e. "some_all" scale includes "none", "some", "most", all")

`m4:` Entailment + empirical alternatives + additional fitted scalar (i.e. "some_all" scale includes "none", "some", "mid", "most", all")

@@

```{r}
performance.output$cor.e6 = as.numeric(performance.output$cor.e6)
performance.output$cor.e11 = as.numeric(performance.output$cor.e11)
performance.output = performance.output %>%
  mutate(avg.corr =
           (cor.e6 + cor.e11) / 2)
performance.output = cbind(performance.output, alphas)
# reorder columns
performance.output = performance.output[, c(1, 2, 3, 5, 4, 6)]
grid.table(performance.output)
```

## Initial thoughts:

### Overall

Across e6 and e11 pragmatic listener judgements we see model fit improve as we add more alternatives. The largest improvement in model fit occurs when we first include the empirically derived salient alternatives (correlations jump +0.23). Have a look at the `m2` vs `m3` correlation for models with pre-RSA normalization below (`c` models):
```{r}
# Tuned, normalized models r
barchart.d = performance.output[c(3, 7, 11, 15), ] %>%
  gather(study, cor, cor.e6, cor.e11)
ggplot(barchart.d, aes(x=model, y=cor, fill=study)) + 
  geom_bar(stat="identity", position="dodge") +
  ylim(0, 1) + 
  ggtitle("Model fit for e6 and e11\npragmatic listener judgments")
  # geom_text(aes(x=model, y=as.numeric(cor) + 0.025, label=round(cor, 3)))
```

In particular, we first start to see a pretty

```{r}
m2.c.plot
m3.c.plot
```

Importantly, we also see improvements with the addition of 'fitted' alternatives in m4 see the plots below:
```{r}
m3.c.plot
m4.c.plot
```

This improvement is particularly clear when we look at e11 data:
```{r}
qplot(stars, listener.p, col=degree, 
      data=m3.c.matched,
      main=paste("m4.c\nalpha: ", alphas[11],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[11, 3]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)

qplot(stars, listener.p, col=degree, 
      data=m4.c.matched,
      main=paste("m4.c\nalpha: ", alphas[15],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[15, 3]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)
```

## Additional notes:

1) In general model performance is greater (correlation) for e11 data. This may be due in part to less extreme literal listener evaluations in e11 (see `RSA+AmbiguousSemantics.Rmd` for a more detailed analysis for literal listener experiment differences). Our best model performance occurs with e11 and the model is m4.c, correlation = `r as.numeric(performance.output[15, 3])` (see plot below):
```{r}
qplot(stars, listener.p, col=degree, 
      data=m4.c.matched,
      main=paste("m4.c\nalpha: ", alphas[15],
                 "\nNormalized = ", T,
                 "\nCorr = ", performance.output[15, 3]),
      ylab="Posterior p(rating | word)") + 
  ylim(0, 1) +
  facet_wrap(~scale) + 
  geom_line(aes(y = pred), lty = 4)
```

@@

3) We may want to explore the the idea of unique alphas for a given scale (would need to change how we've written RSA), for example using scale entropy measures.

# All plots
```{r overall_perf}
m1.a.plot
m1.b.plot
m1.c.plot
m1.d.plot
